---
title: "Step 2 - Probability Occurrance Data"
author: "AmeliaJohnson"
date: "3/25/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Note: This code uses

Spatial_Data_Abridged.xlxs (Marine Scotland): Annual average Surface temperature at each Latitude and Longitude to the 0.1 degree, predicted by our model out 50 years at each decade. 

Herring_Probability_Occurrance.xls (Aquamaps I think): The overall Probability of catching Herring at each Longitude and Latitude to a quarter of a degree. (Not using for population, using this instead.)

Probability Ocurrance

Filter
```{r}
names(Herring_Probability_Occurrance) <- c("Genus", "Species", "Latitude", "Longitude", "C_Square_Code", "Probability")

#subset the UK and surrounding seas
library(dplyr)
Herring_Probability_Occurrance <- Herring_Probability_Occurrance %>% filter(Longitude < 10) %>% filter(Longitude > -15) %>% filter(Latitude < 70) %>% filter(Latitude > 42)
```


Plot Probability Ocurrance
```{r}
library(ggplot2)
ggplot(Herring_Probability_Occurrance, aes(Longitude, Latitude, color = Probability)) + geom_point()
```

Match up Probability Ocurrance Data with Temperature Data:
(Subsetted the Spatial Data to only include the Lat/Long points that we have Probability Data for)

```{r}
library(plyr)
Sub_1 <-match_df(Spatial_Data_Abridged, Herring_Probability_Occurrance, on = c("Latitude", "Longitude"))
Sub_2 <-match_df(Herring_Probability_Occurrance, Spatial_Data_Worst, on = c("Latitude", "Longitude"))

#make Lat and long into one variable
Sub_1 <- Sub_1 %>% mutate(LatLong = (paste(Latitude, Longitude, sep = ", ")))
Sub_2 <- Sub_2 %>% mutate(LatLong = (paste(Latitude, Longitude, sep = ", ")))

#join the two subsets
Sub_A <- join(Sub_1, Sub_2, by = "LatLong", type = "full")

rm(Sub_1)
rm(Sub_2)

```


Plot the temperature of the subset to compare with the population
#omitted (not necessary)
```{r}
#ggplot(Sub, aes(Longitude, Latitude, color = Probability)) + geom_point()
#ggplot(Sub, aes(Longitude, Latitude, color = as.numeric(TempS_Mean))) + geom_point()
```


Find what temperature water the fish are living in now
```{r}
ggplot(data = Sub_A, mapping = aes(x = as.numeric(TempS_Mean) , y = Probability)) + geom_point() + geom_smooth()
ggplot(data = Sub_A, mapping = aes(x = as.numeric(TempS_Mean))) + geom_density()
ggplot(data = Sub_A, mapping = aes(x = Probability)) + geom_density()

```

Run a quadratic model of Temp Vs. Probability at each loction.
```{r}
#fit a -x^2 model
Make_Sub <- Sub_A[(c("Probability", "TempS_Mean"))] %>%
  mutate(
  Temp = (as.numeric(TempS_Mean)), 
  Temp2 = (as.numeric(TempS_Mean)) ^ (2))

square_fit <- lm(Probability ~ Temp + Temp2, Make_Sub)
rm(Make_Sub)
plot(square_fit)
#plot(Sub$Probability, Sub$as.numeric(TempS_Mean), type='l', col='navy', main='Nonlinear relationship', lwd=3)
summary(square_fit)

```

```{r}
rm(Sub_A)
```

