---
title: "Step 2.5 - Baseline"
author: "AmeliaJohnson"
date: "3/25/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Note: This code uses

Spatial_Data_Alternate.xlxs (Marine Scotland): Annual average Surface temperature at each Latitude and Longitude to the 0.1 degree, predicted by the reported 0.11 decgree(C) incrase per decade found in the literature (reference).

Herring_Probability_Occurrance.xls (Aquamaps I think): The overall Probability of catching Herring at each Longitude and Latitude to a quarter of a degree. (Not using for population, using this instead.)

MAP DATA FOR ALTERNATE TEMPERATURE INCREASE RATE: 0.011 C PER YEAR


Map the predicted temperature for each decade deom 2020 to 2070

```{r}
#fix column names
names(Spatial_Data_Alternate) <- c("Latitude", "Longitude", "TempS_Amplitude", "TempS_Mean", "TempS_Max", "X2020", "X2030", "X2040", "X2050", "X2060", "X2070")
```
```{r}


ggplot(Spatial_Data_Alternate, aes(Longitude, Latitude, color = X2020)) + geom_point()
ggplot(Spatial_Data_Alternate, aes(Longitude, Latitude, color = X2030)) + geom_point()
ggplot(Spatial_Data_Alternate, aes(Longitude, Latitude, color = X2040)) + geom_point()
ggplot(Spatial_Data_Alternate, aes(Longitude, Latitude, color = X2050)) + geom_point()
ggplot(Spatial_Data_Alternate, aes(Longitude, Latitude, color = X2060)) + geom_point()
ggplot(Spatial_Data_Alternate, aes(Longitude, Latitude, color = X2070)) + geom_point()

```
We se that the temperature change is very slight.


Match up Probability Ocurrance Data with Temperature Data:
```{r}
library(plyr)
Sub_1 <-match_df(Spatial_Data_Alternate, Herring_Probability_Occurrance, on = c("Latitude", "Longitude"))
Sub_2 <-match_df(Herring_Probability_Occurrance, Spatial_Data_Alternate, on = c("Latitude", "Longitude"))

#make Lat and long into one variable
Sub_1 <- Sub_1 %>% mutate(LatLong = (paste(Latitude, Longitude, sep = ", ")))

Sub_2 <- Sub_2 %>% mutate(LatLong = (paste(Latitude, Longitude, sep = ", ")))

Sub_L <- join(Sub_1, Sub_2, by = "LatLong", type = "full")
```

Use the square_fit model to predict the probability occurrance at each location for each decade

```{r}
library(dplyr)
Sub_L <- Sub_L %>% mutate(Prob_2020 = 0,Prob_2030 = 0,Prob_2040 = 0,Prob_2050 = 0,Prob_2060 = 0,Prob_2070 = 0) 
```

```{r}
#optimal areas
Predict_2020 <- Sub_L[(c("Probability", "X2020"))] %>%
  mutate(
  Temp = X2020, 
  Temp2 = X2020 ^ (2))
Sub_L$Prob_2020 <- predict(square_fit, Predict_2020)
rm(Predict_2020)

Predict_2030 <- Sub_L[(c("Probability", "X2030"))] %>%
  mutate(
  Temp = X2030, 
  Temp2 = X2030 ^ (2))
Sub_L$Prob_2030 <- predict(square_fit, Predict_2030)
rm(Predict_2030)

Predict_2040 <- Sub_L[(c("Probability", "X2040"))] %>%
  mutate(
  Temp = X2040, 
  Temp2 = X2040 ^ (2))
Sub_L$Prob_2040 <- predict(square_fit, Predict_2040)
rm(Predict_2040)

Predict_2050 <- Sub_L[(c("Probability", "X2050"))] %>%
  mutate(
  Temp = X2050, 
  Temp2 = X2050 ^ (2))
Sub_L$Prob_2050 <- predict(square_fit, Predict_2050)
rm(Predict_2050)

Predict_2060 <- Sub_L[(c("Probability", "X2060"))] %>%
  mutate(
  Temp = X2060, 
  Temp2 = X2060 ^ (2))
Sub_L$Prob_2060 <- predict(square_fit, Predict_2060)
rm(Predict_2060)

Predict_2070 <- Sub_L[(c("Probability", "X2070"))] %>%
  mutate(
  Temp = X2070, 
  Temp2 = X2070 ^ (2))
Sub_L$Prob_2070 <- predict(square_fit, Predict_2070)
rm(Predict_2070)

```

replace all negative predictions with 0 and all predictions > 1 with 1 (Probability must be between 0 and 1)
```{r}
Sub_L$Prob_2020[Sub_L$Prob_2020 < 0] = 0
Sub_L$Prob_2030[Sub_L$Prob_2030 < 0] = 0
Sub_L$Prob_2040[Sub_L$Prob_2040 < 0] = 0
Sub_L$Prob_2050[Sub_L$Prob_2050 < 0] = 0
Sub_L$Prob_2060[Sub_L$Prob_2060 < 0] = 0
Sub_L$Prob_2070[Sub_L$Prob_2070 < 0] = 0

```



```{r}
ggplot(Sub_L, aes(Longitude, Latitude, color = Prob_2020)) + geom_point() + scale_color_gradient(low = "red", high = "blue")

ggplot(Sub_L, aes(Longitude, Latitude, color = Prob_2030)) + geom_point()+ scale_color_gradient(low = "red", high = "blue")

ggplot(Sub_L, aes(Longitude, Latitude, color = Prob_2040)) + geom_point()+ scale_color_gradient(low = "red", high = "blue")

ggplot(Sub_L, aes(Longitude, Latitude, color = Prob_2050)) + geom_point()+ scale_color_gradient(low = "red", high = "blue")

ggplot(Sub_L, aes(Longitude, Latitude, color = Prob_2060)) + geom_point()+ scale_color_gradient(low = "red", high = "blue")

ggplot(Sub_L, aes(Longitude, Latitude, color = Prob_2070)) + geom_point()+ scale_color_gradient(low = "red", high = "blue")

```

We see that over 50 years, the Herring population is barely mocing at all. We know this is not accurate because studies show that fish populations are already moving. (reference)