---
title: "Step 1 - find Beta and confint"
author: "AmeliaJohnson"
date: "3/25/2020"---
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Note: This code uses

Monthly_Data.cvs (Marine Scotland): Records all the monthly surface areas for 13 Scottish Sea Areas from 1997 to 2014. (No longitude and Latitude)



Look at the monthly temperature data for 1985 - 2014
```{r}
names(Monthly_Data)
library(ggplot2)
ggplot(data = Monthly_Data, mapping = aes(x = Decimal_Year , y = Monthly_Temperature)) + geom_line() + geom_smooth(method = "lm")

ggplot(data = Monthly_Data, mapping = aes(x = Decimal_Year , y = Monthly_Temperature, color = Sea_Area)) + geom_line()  + geom_smooth(method = "lm", se = F)

ggplot(data = Monthly_Data, mapping = aes(x = Decimal_Year , y = Monthly_Temperature, color = Sea_Area)) + geom_line() + geom_smooth(se = F) 
```

Run a regression on the historical temperature data to find a Beta for monthly temperature increase

Monthly_Temp = Beta_0 + Beta_1(Decimal_Year)

interpretations: For each month increase, the Temp increases by Beta_1
```{r}
lm_fit <- lm(Monthly_Temperature ~ Decimal_Year, data = Monthly_Data)

summary(lm_fit)
confint(lm_fit)
confint(lm_fit, level = 0.80)

plot(lm_fit)

ggplot(Monthly_Data, aes(Monthly_Temperature)) + geom_density()


```
Beta_1: 0.015137
95% confidence intervals for the Betas: (-0.002244637,  0.03251833)
80% confidence intervals for the Betas: (0.003774064, 0.02649963)

We now make an excel spreadsheet to predict the temperature over all the decades from 2020-2070 based on Beta_1 = 0.015137. (We will use the 80% confidence interval later because we know that the Beta is not negative.) 

For these predicted values, we use the formula: 

Not a regression: Temp_future = Temp_2014 + (0.015137*12)(Year - 2014)

(We assume Temp_2014 =TempS_Mean)


We will use the annual mean temperature to make these excel documents. The data are pictured here:
Map the mean temperature
```{r}
#plot the map
ggplot(Spatial_Data_Abridged, aes(Longitude, Latitude, color = as.numeric(TempS_Mean))) + geom_point()
```
