---
title: "Step 3 - Decade, mean Beta"
author: "AmeliaJohnson"
date: "3/25/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Note: This code uses

Spatial_Data_Abridged.xlxs (Marine Scotland): Annual average Surface temperature at each Latitude and Longitude to the 0.1 degree, predicted by our model out 50 years at each decade. 

Herring_Probability_Occurrance.xls (Aquamaps I think): The overall Probability of catching Herring at each Longitude and Latitude to a quarter of a degree. (Not using for population, using this instead.)



Map the predicted  temperature for each decade deom 2020 to 2070 based on the Beta from the Monthly Data linear regression

```{r}
#fix column names
names(Spatial_Data_Abridged) <- c("Latitude", "Longitude", "TempS_Amplitude", "TempS_Mean", "TempS_Max", "X2020", "X2030", "X2040", "X2050", "X2060", "X2070", "X2021", "X2022", "X2023", "X2024", "X2025", "X2026", "X2027", "X2028", "X2029")
```
```{r}
Spatial_Data_Abridged <- Spatial_Data_Abridged[-1,]

ggplot(Spatial_Data_Abridged, aes(Longitude, Latitude, color = X2020)) + geom_point()
ggplot(Spatial_Data_Abridged, aes(Longitude, Latitude, color = X2030)) + geom_point()
ggplot(Spatial_Data_Abridged, aes(Longitude, Latitude, color = X2040)) + geom_point()
ggplot(Spatial_Data_Abridged, aes(Longitude, Latitude, color = X2050)) + geom_point()
ggplot(Spatial_Data_Abridged, aes(Longitude, Latitude, color = X2060)) + geom_point()
ggplot(Spatial_Data_Abridged, aes(Longitude, Latitude, color = X2070)) + geom_point()

```

Match up Probability Ocurrance Data with Temperature Data:
(Sub_Msetted the Spatial Data to only include the Lat/Long points that we have Probability Data for)

```{r}
library(plyr)
Sub_M_1 <-match_df(Spatial_Data_Abridged, Herring_Probability_Occurrance, on = c("Latitude", "Longitude"))
Sub_M_2 <-match_df(Herring_Probability_Occurrance, Spatial_Data_Worst, on = c("Latitude", "Longitude"))

#make Lat and long into one variable
Sub_M_1 <- Sub_M_1 %>% mutate(LatLong = (paste(Latitude, Longitude, sep = ", ")))
Sub_M_2 <- Sub_M_2 %>% mutate(LatLong = (paste(Latitude, Longitude, sep = ", ")))

#join the two Sub_Msets
Sub_M <- join(Sub_M_1, Sub_M_2, by = "LatLong", type = "full")

rm(Sub_M_1)
rm(Sub_M_2)

```


Use the square_fit model to predict the probability occurrance at each location for each decade

```{r}
library(dplyr)
Sub_M <- Sub_M %>% mutate(Prob_2020 = 0,Prob_2030 = 0,Prob_2040 = 0,Prob_2050 = 0,Prob_2060 = 0,Prob_2070 = 0) 
```

```{r}
#optimal areas
Predict_2020 <- Sub_M[(c("Probability", "X2020"))] %>%
  mutate(
  Temp = X2020, 
  Temp2 = X2020 ^ (2))
Sub_M$Prob_2020 <- predict(square_fit, Predict_2020)
rm(Predict_2020)

Predict_2030 <- Sub_M[(c("Probability", "X2030"))] %>%
  mutate(
  Temp = X2030, 
  Temp2 = X2030 ^ (2))
Sub_M$Prob_2030 <- predict(square_fit, Predict_2030)
rm(Predict_2030)

Predict_2040 <- Sub_M[(c("Probability", "X2040"))] %>%
  mutate(
  Temp = X2040, 
  Temp2 = X2040 ^ (2))
Sub_M$Prob_2040 <- predict(square_fit, Predict_2040)
rm(Predict_2040)

Predict_2050 <- Sub_M[(c("Probability", "X2050"))] %>%
  mutate(
  Temp = X2050, 
  Temp2 = X2050 ^ (2))
Sub_M$Prob_2050 <- predict(square_fit, Predict_2050)
rm(Predict_2050)

Predict_2060 <- Sub_M[(c("Probability", "X2060"))] %>%
  mutate(
  Temp = X2060, 
  Temp2 = X2060 ^ (2))
Sub_M$Prob_2060 <- predict(square_fit, Predict_2060)
rm(Predict_2060)

Predict_2070 <- Sub_M[(c("Probability", "X2070"))] %>%
  mutate(
  Temp = X2070, 
  Temp2 = X2070 ^ (2))
Sub_M$Prob_2070 <- predict(square_fit, Predict_2070)
rm(Predict_2070)

```

replace all negative predictions with 0 and all predictions > 1 with 1 (Probability must be between 0 and 1)
```{r}
Sub_M$Prob_2020[Sub_M$Prob_2020 < 0] = 0
Sub_M$Prob_2030[Sub_M$Prob_2030 < 0] = 0
Sub_M$Prob_2040[Sub_M$Prob_2040 < 0] = 0
Sub_M$Prob_2050[Sub_M$Prob_2050 < 0] = 0
Sub_M$Prob_2060[Sub_M$Prob_2060 < 0] = 0
Sub_M$Prob_2070[Sub_M$Prob_2070 < 0] = 0

```



```{r}
ggplot(Sub_M, aes(Longitude, Latitude, color = Prob_2020)) + geom_point() + scale_color_gradient(low = "red", high = "blue")

ggplot(Sub_M, aes(Longitude, Latitude, color = Prob_2030)) + geom_point()+ scale_color_gradient(low = "red", high = "blue")

ggplot(Sub_M, aes(Longitude, Latitude, color = Prob_2040)) + geom_point()+ scale_color_gradient(low = "red", high = "blue")

ggplot(Sub_M, aes(Longitude, Latitude, color = Prob_2050)) + geom_point()+ scale_color_gradient(low = "red", high = "blue")

ggplot(Sub_M, aes(Longitude, Latitude, color = Prob_2060)) + geom_point()+ scale_color_gradient(low = "red", high = "blue")

ggplot(Sub_M, aes(Longitude, Latitude, color = Prob_2070)) + geom_point()+ scale_color_gradient(low = "red", high = "blue")

```

We see that by 2040, the probability occurrance is 0 for all areas around Scotland. The fish will move completely before this time, and will be mostly gone by 2030. 

We add the columns of the years between 2020 and 2030 to the data set.

Use the square_fit model to predict the probability occurrance at each locaton for each year between 2020 and 2030.

```{r}
library(dplyr)
Sub_M <- Sub_M %>% mutate(Prob_2021 = 0,Prob_2022 = 0,Prob_2023 = 0,Prob_2024 = 0,Prob_2025 = 0,Prob_2026 = 0,Prob_2027 = 0,Prob_2028 = 0,Prob_2029 = 0) 
```

```{r}
#optimal areas
Predict_2021 <- Sub_M[(c("Probability", "X2021"))] %>%
  mutate(
  Temp = X2021, 
  Temp2 = X2021 ^ (2))
Sub_M$Prob_2021 <- predict(square_fit, Predict_2021)
rm(Predict_2021)

Predict_2022 <- Sub_M[(c("Probability", "X2022"))] %>%
  mutate(
  Temp = X2022, 
  Temp2 = X2022 ^ (2))
Sub_M$Prob_2022 <- predict(square_fit, Predict_2022)
rm(Predict_2022)

Predict_2023 <- Sub_M[(c("Probability", "X2023"))] %>%
  mutate(
  Temp = X2023, 
  Temp2 = X2023 ^ (2))
Sub_M$Prob_2023 <- predict(square_fit, Predict_2023)
rm(Predict_2023)

Predict_2024 <- Sub_M[(c("Probability", "X2024"))] %>%
  mutate(
  Temp = X2024, 
  Temp2 = X2024 ^ (2))
Sub_M$Prob_2024 <- predict(square_fit, Predict_2024)
rm(Predict_2024)

Predict_2025 <- Sub_M[(c("Probability", "X2025"))] %>%
  mutate(
  Temp = X2025, 
  Temp2 = X2025 ^ (2))
Sub_M$Prob_2025 <- predict(square_fit, Predict_2025)
rm(Predict_2025)

Predict_2026 <- Sub_M[(c("Probability", "X2026"))] %>%
  mutate(
  Temp = X2026, 
  Temp2 = X2026 ^ (2))
Sub_M$Prob_2026 <- predict(square_fit, Predict_2026)
rm(Predict_2026)

Predict_2027 <- Sub_M[(c("Probability", "X2027"))] %>%
  mutate(
  Temp = X2027, 
  Temp2 = X2027 ^ (2))
Sub_M$Prob_2027 <- predict(square_fit, Predict_2027)
rm(Predict_2027)

Predict_2028 <- Sub_M[(c("Probability", "X2028"))] %>%
  mutate(
  Temp = X2028, 
  Temp2 = X2028 ^ (2))
Sub_M$Prob_2028 <- predict(square_fit, Predict_2028)
rm(Predict_2028)

Predict_2029 <- Sub_M[(c("Probability", "X2029"))] %>%
  mutate(
  Temp = X2029, 
  Temp2 = X2029 ^ (2))
Sub_M$Prob_2029 <- predict(square_fit, Predict_2029)
rm(Predict_2029)



```

replace all negative predictions with 0 and all predictions > 1 with 1 (Probability must be between 0 and 1)
```{r}
Sub_M$Prob_2021[Sub_M$Prob_2021 < 0] = 0
Sub_M$Prob_2022[Sub_M$Prob_2022 < 0] = 0
Sub_M$Prob_2023[Sub_M$Prob_2023 < 0] = 0
Sub_M$Prob_2024[Sub_M$Prob_2024 < 0] = 0
Sub_M$Prob_2025[Sub_M$Prob_2025 < 0] = 0
Sub_M$Prob_2026[Sub_M$Prob_2026 < 0] = 0
Sub_M$Prob_2027[Sub_M$Prob_2027 < 0] = 0
Sub_M$Prob_2028[Sub_M$Prob_2028 < 0] = 0
Sub_M$Prob_2029[Sub_M$Prob_2029 < 0] = 0


```



```{r}
ggplot(Sub_M, aes(Longitude, Latitude, color = Prob_2021)) + geom_point() + scale_color_gradient(low = "red", high = "blue")
ggplot(Sub_M, aes(Longitude, Latitude, color = Prob_2022)) + geom_point() + scale_color_gradient(low = "red", high = "blue")
ggplot(Sub_M, aes(Longitude, Latitude, color = Prob_2023)) + geom_point() + scale_color_gradient(low = "red", high = "blue")
ggplot(Sub_M, aes(Longitude, Latitude, color = Prob_2024)) + geom_point() + scale_color_gradient(low = "red", high = "blue")
ggplot(Sub_M, aes(Longitude, Latitude, color = Prob_2025)) + geom_point() + scale_color_gradient(low = "red", high = "blue")
ggplot(Sub_M, aes(Longitude, Latitude, color = Prob_2026)) + geom_point() + scale_color_gradient(low = "red", high = "blue")
ggplot(Sub_M, aes(Longitude, Latitude, color = Prob_2027)) + geom_point() + scale_color_gradient(low = "red", high = "blue")
ggplot(Sub_M, aes(Longitude, Latitude, color = Prob_2028)) + geom_point() + scale_color_gradient(low = "red", high = "blue")
ggplot(Sub_M, aes(Longitude, Latitude, color = Prob_2029)) + geom_point() + scale_color_gradient(low = "red", high = "blue")


```

The mean scenario shows that the fish occurance probability will be very low by around 2029. 